Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-08 src/ms.tex
Updating tex file

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
%\revised{\today}
%\accepted{\today}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{The Rossby Ridge}
\shortauthors{David et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
%% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{figures/}}
%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

% \title{The Rossby Ridge: a Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages}
%\title{A Pile-up of Rapidly Rotating Main-Sequence Stars with a Broad Range of Ages: \\Further Evidence for Weakened Magnetic Braking}
%\title{Two Pile-ups in the Temperature--Period Distribution of Main-Sequence Stars:\\Possible Evidence for Weakened Magnetic Braking and Core-Envelope Coupling}
\title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}

@@

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin. Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin. Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution. {\color{blue}The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters. Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.}
\end{abstract}

@@

%Nevertheless, it is by now clear that the simple $t^{1/2}$ scaling of rotation rates with age breaks down in certain regimes of stellar mass and age. Efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}. More recently, low-mass stars were shown to experience a temporary epoch of stalled spin-down which is not accounted for by standard gyrochronology models \citep{Curtis2019a}. \citet{Spada2020} and \citet{Curtis2020} proposed that this epoch of apparent stalling may arise from rotational coupling of the core and envelope, in which redistribution of angular momentum from the interior to the surface temporarily offsets the effects of magnetic braking. Regardless of the true mechanism, rotation periods in this part of parameter space have diminished utility in predicting stellar ages.

@@

%We argue that several aspects of the long-period pile-up feature are consistent with expectations from the WMB model. The short-period pile-up, by comparison, may be due a temporary epoch of stalled wind-driven losses due to core-envelope coupling.

@@

\subsection{Initial observations}
A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}. Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}). {\color{blue} A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample. As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.}

@@

In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}. A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling. {\color{blue}This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.}

@@

While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz. {\color{blue} We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.}

{\color{blue} Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models. In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e. the low-mass stars would be rotating more slowly than the model predictions). Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$. A higher normalization factor would cause the low-mass stars to spin down more at fixed age.}

{\color{blue} Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015}. When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.}

@@

% \subsection{Piecewise polynomial model}
% We fit a piecewise linear relation to the CKS stars on the ridge with $5800~K<$~\teff~$<6400~K$. We performed an initial Levenberg-Marquardt least-squares fit with the \texttt{curve\_fit} function in the \texttt{scipy.optimize} class. We then performed Markov chain Monte Carlo sampling (MCMC) with the \texttt{emcee} package \citep{emcee2013, emcee2019}.

@@

%{\color{blue} From a sample of Kepler targets with asteroseismic rotation rates, \citet{Hall2021} found a good match to observations when adopting a critical Rossby number of $\mathrm{Ro_{crit}} = 1.97$.}

@@

%We computed constant Rossby models across the range of \teff bins using the equation for the convective turnover timescale presented in \citet{Gunn1998} and replicated in \citet{CranmerSaar2011}. This equation is valid in the \teff range of 3300~K$\lesssim \teff \lesssim$~7000~K. %\textbf{Note: we can also try Landin 2010 and Barnes Kim 2010 formulas for convective turnover timescales.}

@@

{\color{blue}  We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set. For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d. To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$. We note that this selection is particular to the CKS \teff scale and is not general. For the CKS sample, we additionally required $\logg~>~4$ and 5850~K~$<~\teff~<$~6500~K. In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.}

@@

{\color{blue} We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation. Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 may be compatible with all of the data sets if modest \teff systematics are present. Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff})}.

@@

{\color{blue} We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex. While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}). Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pile-up stars. We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.}

@@

\subsection{Where do the pile-ups end?} \label{subsec:extent}
{\color{blue} It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff. While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach. We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample. In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window. We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, approximately halving in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}). The relative fraction of stars on the short-period edge declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal. Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.}

@@

\subsection{Long-period behavior} \label{subsec:longperiod}
{\color{blue}
As discussed in \S\ref{subsec:models} and \S\ref{subsec:rossby}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}. The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021} and we suggest the two features are one and the same. Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.

Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model. Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr. By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr. {\color{blue} The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz. We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.}
}

\subsection{Short-period behavior} \label{subsec:shortperiod}
{\color{blue}
Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model. However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one. Works examining the rotation sequences of open clusters have found overlap between low-mass members of clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}. In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.

@@

\textbf{We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}. The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct. As mentioned above, the leading theory for the earlier stage of stalled braking is core-envelope coupling. Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core-envelope coupling. In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia. The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.}
}

\subsection{Implications for the period gap}
{\color{blue}
An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for stars of similar \teff. The effect was first noticed for M-dwarfs \citep{McQuillan2013b}, but was later shown to extend to $\sim$5000~K \citep{Reinhold2013, McQuillan2014, ReinholdHekker2020}, and even to $\sim$6500~K \citep{Davenport2017}. \citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}. However, \citet{Reinhold2019} found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability. Consequently, those authors proposed that the period bimodality may be the result of a transition between spot- and faculae-dominated photospheres. In this scenario, the period gap is due to bright faculae canceling out the effects of dark star spots.

The short- and long-period pile-ups we examine here naturally produce a dearth of rotators at intermediate periods. This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \citet{McQuillan2014} sample. Moreover, as seen in Figure~\ref{fig:gap}, we recover the gradient in photometric variability across the gap pointed out by \citet{Reinhold2019}. The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum at the location of the gap. This supports the notion that the cause of the gap is due to changes in the stars themselves, rather than being the result of mixed stellar populations. However, the variability levels on both sides of the gap are not close to the detection limit, as evidenced by the fact that periods are securely detected for stars with similar properties at much lower \rper values. This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap. 
}

@@

{\color{blue}
It is unclear whether or not the Sun is a resident of the long-period pile-up. If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$~5--7 days above the long-period pile-up. This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase. However, a systematic offset in \teff could place the Sun on the long-period pile-up. Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up. Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required. To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K. However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.

@@

Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}). If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out. One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021}. Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question. 
}

@@

    \item {\color{blue} It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it. Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up. If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.}

    %\textbf{Change this:} {\color{blue}However, this statement depends on (i) the magnitude of any shift between the \teff scales of the Sun and the spectroscopic surveys used in this work, as well as (ii) any shift between the age scales of the solar system and the theoretical isochrones used to age-date the CKS sample. Stars which populate the ridge with isochrone ages older than the Sun may in fact have ages younger than or equal to the Sun's if systematic offsets exist.}

>>>>

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
\revised{\today}
%\accepted{\today}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{The Rossby Ridge}
\shortauthors{David et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
%% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%% \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{figures/}}

\begin{document}
\title{Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution}

@@

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST--Kepler survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin. Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin. Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the rotation period distribution. The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of delayed spin-down due to core-envelope coupling previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters. Together, the short- and long-period pile-ups create an underdensity of intermediate rotation periods which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars.
\end{abstract}

@@

@@

@@

\subsection{Initial observations}
A pile-up of stars along the long-period edge for stars with \teff~$>5800$~K was first noted when examining the \teff-\prot plane for the CKS sample, using the CKS \teff values and the vetted rotation period compilation from \citet{David2021}. Sourcing rotation periods from \citet{McQuillan2013, Mazeh2015} and \citet{Angus2018} revealed that this pile-up is still apparent when adopting periods uniformly from a single catalog (Figure~\ref{fig:ridge}). A secondary pile-up at the short-period edge is also apparent, though less pronounced, in the \teff--\prot distribution of the \lamostmcq sample. As shown in Figure~\ref{fig:kde}, This secondary pile-up is seen most clearly through applying Gaussian kernel density estimation (KDE), which was performed with the \texttt{seaborn} package \citep{seaborn}.

@@

In both the CKS and LAMOST samples, there is a change of slope along the ridge, with an inflection point corresponding closely to the Kraft break at $\approx$~6250~K, the point at which convective envelopes become vanishingly thin \citep{Kraft1967}. A piecewise linear fit to the ridge in the CKS sample confirmed that the inflection point occurs at \teff~=~6250~$\pm$~32~K, where the uncertainty was estimated from Markov chain Monte Carlo (MCMC) sampling. This change in slope is likely due to the fact that the convective turnover timescale, \taucz, changes rapidly above 6250~K.

@@

While better agreement between the WMB model and observations might be achieved with stalling at a lower \rocrit, it is also possible that there are systematic offsets in the \teff scales between the observations and models used in \citet{vanSaders2019}, as well as differences in the computation of \taucz.  We also note that, while the models were computed using a simulated Kepler stellar population and selection function, the actual observed population and selection function of the LAMOST--Kepler catalog may be slightly different.

Shifting the LAMOST \teff to higher values would bring the data into better agreement with the models. In turn, it appears that the long-period edge for lower mass stars would be at higher \prot than the models (i.e. the low-mass stars would be rotating more slowly than the model predictions). Such a discrepancy could result from different underlying populations between the models and the \lamostmcq sample, or a different normalization for the magnetic braking law; the models above employ a modified magnetic braking law that is scaled to match the rotation period of the Sun \citep[see equations 1 \& 2 of][]{vanSaders2013}, with a normalization factor of $f_K = 6.6$. A higher normalization factor would cause the low-mass stars to spin down more at fixed age.

Notably, both models fail to reproduce the observed short-period edge, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015}. When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.

@@

@@

@@

@@

We additionally fit the CKS long-period pile-up (using three different homogeneous \teff sources), the \citet{Hall2021} asteroseismic main-sequence sample, and all combined data sets of the long-period pile-up, allowing for \teff offsets in each data set. For the CKS pile-up stars, the \prot uncertainties were assumed to be 1~d. To isolate the long-period pile-up stars in the CKS sample, we selected a trapezoidal region using the condition $-0.0314~\teff + 199.286 < \prot < -0.0314~\teff + 206.286$. We note that this selection is particular to the CKS \teff scale and is not general. For the CKS sample, we additionally required $\logg~>~4$ and 5850~K~$<~\teff~<$~6500~K. In the joint fit of all data sets, the log likelihood was defined as the sum of log likelihoods (given by Eq.~\ref{eq:3}) from the individual data sets.

@@

We caution that the specific values of \rocrit found here are highly sensitive to shifts in the \teff scales between the models and observations, as well as the \teff scale of the models used to derive the \taucz relation. Therefore, while we find values of \rocritfinal are able to explain the various individual data sets, a value of \rocrit  $\approx$~2 may be compatible with all of the data sets if modest \teff systematics are present. Comparison of the long-period pile-ups as seen in the CKS and \lamostmcq samples support the notion that the LAMOST \teff scale is about 30--40~K cooler than the CKS scale (see also Appendix~\ref{app:teff}).

@@

We additionally selected stars on the long-period pile-up from the \lamostmcq sample by selecting stars with periods within 5\% of the Ro~=~1.3 curve (which traces the center of the highest density contour in Figure~\ref{fig:kde}) , 5500~K~$<$~\teff~$<$~6500~K, and 4~dex~$<$~\logg~$<$~4.75~dex. While ages for the \lamostkep sample are not available, the broad distribution of these stars in the spectroscopic H-R diagram supports the inference from the CKS sample that the long-period pile-up is populated by stars with a broad range of ages (Figure~\ref{fig:ages}). Interestingly, the solar \teff and \logg values appear to be wholly consistent with the distribution of long-period pile-up stars. We discuss the Sun in context of the long-period pile-up further in \S\ref{subsec:thesun}.

@@

\subsection{Where do the pile-ups end?} \label{subsec:extent}
It appears from Figures~\ref{fig:kde} and \ref{fig:models} that the number density of stars on the long- and short-period pile-ups declines towards cooler \teff. While it not clear whether or not these declines are astrophysical in nature, the result of the Kepler selection function or observational biases, or some combination of effects, we attempted to characterize the extent of the pile-ups through the following approach. We found through inspection that constant Rossby curves of Ro=0.5 and Ro=1.3 (for the short- and long-period pile-ups, respectively) appear to describe the highest density contours found from Gaussian kernel density estimation of the \lamostmcq sample. In windows of 10~K width we measured the fraction of stars with periods within 1~d of each of the two constant Rossby curves, relative to the total number of stars in that \teff window. We found that the relative fraction of stars on the long-period pile-up declines rapidly between 6500~K and 5800~K, by more than half in that temperature range before plateauing at cooler \teff (Figure~\ref{fig:fraction}). The relative fraction of stars on the short-period edge declines more slowly, and below the temperature of the Sun, the relative fractions of stars on the two pile-ups are nearly equal. Thus, it is not yet clear if either pile-up extends to temperatures cooler than \tsun.

@@

\subsection{Long-period behavior} \label{subsec:longperiod}

As discussed in \S\ref{subsec:models} and \S\ref{subsec:rossby}, the existence of a pile-up of stars with a constant Rossby number is consistent with expectations from the WMB model of \citet{vanSaders2016, vanSaders2019}. The location of the long-period pile-up coincides closely with the cluster of main-sequence stars with asteroseismic rotation rates determined by \citet{Hall2021} and we suggest the two features are one and the same. Offsets in both observational and theoretical \teff scales between different studies can explain the modest discrepancies between the asteroseismic sample and the samples considered here, as well as discrepancies in the \rocrit values inferred here and those proposed in \citet{vanSaders2019}.

Recently, \citet{LorenzoOliveira2019} studied the \prot--age relation of solar twins observed by the Kepler mission, finding marginal statistical evidence in favor of a standard spin-down model over the WMB model. Those authors posited that if WMB takes place for Sun-like stars, it should happen at \rocrit~$\gtrsim 2.29$ or ages~$\gtrsim$~5.3~Gyr. By comparison, our findings provide support for the WMB model among stars that are slightly hotter than the Sun, but at \rocrit~$\lesssim$~2 and at ages in the range of $\sim$2--6~Gyr. The \rocrit values we found in \S\ref{subsec:rossby} are somewhat smaller than the values of $\sim$2 found by \citet{vanSaders2016,vanSaders2019}, however this is sensitive to different \teff scales and different prescriptions for \taucz. We explored the \taucz formulae of \citet{BarnesKim2010} and \citet{Landin2010}, but these produce larger \taucz at fixed \teff than the \citet{Gunn1998} relation, which exacerbates the \rocrit mismatch mentioned above.

\subsection{Short-period behavior} \label{subsec:shortperiod}

Unlike the long-period pile-up, the short-period pile-up is not predicted by the WMB model. However, this feature may also be due to an epoch of stalled wind-driven angular momentum losses, albeit a temporary one. Works examining the rotation sequences of open clusters have found overlap between low-mass members of clusters of different ages \citep{Agueros2018, Curtis2019a, Curtis2020}. In other words, the spin rates of low-mass stars appear to evolve very little in the time elapsed between the ages of those clusters.

@@

We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}. The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct. As mentioned above, the leading theory for the earlier stage of stalled braking is core-envelope coupling. Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core-envelope coupling. In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia. The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.

\subsection{Implications for the period gap} \label{subsec:gap}

An unexplained feature of the Kepler rotation period distribution is the existence of a bimodal period distribution for stars of similar \teff. The effect was first noticed for M-dwarfs \citep{McQuillan2013b}, but was later shown to extend to $\sim$5000~K \citep{Reinhold2013, McQuillan2014, ReinholdHekker2020}, and even to $\sim$6500~K \citep{Davenport2017}. \citet{McQuillan2013b, McQuillan2014} speculated that this period bimodality could originate from stellar populations of different ages, an explanation seemingly supported by a correlation between the strength of the bimodality and height above the galactic disk \citep{DavenportCovey2018}. However, \citet{Reinhold2019} found that the dearth of stars with intermediate rotation periods is associated with a decrease in photometric variability. Consequently, those authors proposed that the period bimodality may be the result of a transition between spot- and faculae-dominated photospheres. In this scenario, the period gap is due to bright faculae canceling out the effects of dark star spots.

The short- and long-period pile-ups we examine here naturally produce a dearth of rotators at intermediate periods. This gap is the same period gap noticed by the authors mentioned above, as made apparent when comparing the \lamostmcq sample to the original \citet{McQuillan2014} sample. Moreover, as seen in Figure~\ref{fig:gap}, we recover the gradient in photometric variability across the gap pointed out by \citet{Reinhold2019}. The photometric variability, as measured through the \rper metric published by \citet{McQuillan2014}, reaches a local minimum at the location of the gap. This supports the notion that the cause of the gap is due to changes in the stars themselves, rather than being the result of mixed stellar populations. However, the variability levels on both sides of the gap are not close to the detection limit, as evidenced by the fact that periods are securely detected for stars with similar properties at much lower \rper values. This would suggest that the period gap is not due solely to a detection issue, unless variability levels were to drop precipitously as stars approached the gap.

@@

It is unclear whether or not the Sun is a resident of the long-period pile-up. If one assumes there are no systematic offsets between the \teff scales of the Sun and the spectroscopic catalogs considered here, the Sun's equatorial rotation period places it $\sim$~5--7 days above the long-period pile-up. This raises the question of how the Sun's angular momentum has evolved to its current state, and whether WMB is a generic evolutionary phase. However, a systematic offset in \teff could place the Sun on the long-period pile-up. Shifting the CKS \teff higher by $\sim$50--100~K would appear to place the Sun closer to the pile-up. Using the SPOCS \teff for the CKS sample, a shift of $\sim$100--200~K would be required. To place the Sun on the long-period pile-up in the \lamostmcq sample would require shifting the LAMOST \teff higher by $\sim$200--300~K. However, comparing the Sun to the \citet{Hall2021} asteroseismic sample (Figure~\ref{fig:kde}) seems to support the notion that the Sun may indeed reside on the long-period pile-up, or very close to it, without the need for a \teff shift.

@@

Another question posed by the Sun in relation to the long-period pile-up is how such a feature can appear so well-defined given the spread of periods observed on the Sun due to differential rotation (Figure~\ref{fig:surveys}). If other Sun-like stars have a similar degree of differential rotation, one might expect the long-period edge to be blurred out. One possible explanation is that starspots at higher latitudes produce smaller variations in light curves as they are in view or out of view for most of a rotation cycle for some random inclination \citep[e.g.][]{Luger2021}. Forward modeling simulations of the Kepler field including the effects of surface differential rotation may be able to answer this question.

@@

    \item It is yet unclear whether the Sun resides on the long-period pile-up or has already evolved through it. Shifts of $\sim$50--300~K between the \teff scales of the Sun and the spectroscopic surveys studied here could place the Sun on the long-period pile-up. If the Sun has evolved through the pile-up, there is some modest tension with the Sun's age and the ages of the oldest stars on the pile-up, which might be explained by inaccurate isochrone ages.

    \item We observe a secondary overdensity of stars at the short-period edge of the \teff-\prot plane. This overdensity is less prominent than the long-period pile-up, potentially due to the age distribution of the Kepler field or because this pile-up is shorter-lived relative to the long-period pile-up. The short-period pile-up appears to be intermediate to the empirical Pleiades (0.12~Gyr) and Praesepe (0.67~Gyr) open cluster sequences and may result from a temporary epoch of stalled spin-down due to core-envelope coupling, as proposed by \citet{Curtis2020}. The short-period pile-up can also be fit with a constant Rossby model, though over a range of \teff that differs from that of the long-period pile-up.

    \item The number density of stars on the long-period pile-up declines with \teff, though it remains unclear whether this is due to astrophysics, the Kepler selection function, observational biases, or some combination of effects. The relative fraction of stars on the long-period pile up declines by a factor of $\sim$2 between $\sim$6200~K and $\sim$5800~K.
